---
title: "EEB313 - Project Trial 2"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

#### ctrl/cmd+F "CODE" to go straight to the code

It's a mess I recommend going to "Visual" mode for non-code stuff

## Parameters

beta_0 - Transmission rate of unvaccinated\
beta_A - Transmission rate of A vaccine\
beta_B - Transmission rate of B vaccine\
beta_AB - Transmission rate of AB vaccine\

gamma_0 - Recovery rate of unvaccinated\
gamma_A - Recovery rate for vaccine A\
gamma_B - Recovery rate for vaccine B\
gamma_AB - Recovery rate for both vaccines\

c - Rate of culling

V_A - vaccine rate of vaccine A\
V_B - vaccine rate of vaccine B\
V_AB - rate of getting vaccinated by both A and B\

## Differential Equations

S - Unvaccinated and susceptible S_A - vaccinated with A and susceptible
S_B - vaccinated with B and susceptible S_AB - vaccinated with AB and
susceptible

I - Infected and unvaccianted I_A - Infected and vaccinated with A I_B -
Infected and vaccinated with B I_AB - Infected and vaccinated with AB

R - Recovery of unvaccinated individuals R_A - Recovery of vaccinated A
individuals R_B - Recovery of vaccinated A individuals R_AB - Recovery
of vaccinated A individuals

## Equations

Let us know if we missed any factors or need to remove some!

### Susceptible

$$
  dS/dt = -S*(\beta_A*I_A + \beta_B*I_B + \beta_AB*I_{AB} + \beta_0*I + V_A + V_B + V_{AB})
$$

Breakdown:\
$-S\beta_A*I_A$ -\> Infection by individual with vaccine A\
$-S\beta_B*I_B$ -\> Infection by individual with vaccine B\
$-S\beta_AB*I_AB$ -\> Infection by individual with vaccine AB\
$-S\beta_0*I$ -\> Infection by individual that's unvaccinated\
$-S*V_A$ -\> Remove individuals that get vaccinated with vaccine A\
$-S*V_A$ -\> Remove individuals that get vaccinated with vaccine B\
$-S*V_AB$ -\> Remove individuals that get vaccinated with both vaccine A
& B

*This logic is similar to* $dS_A/dt$, $dS_B/dt$, $dS_AB$ *Only
differences will be explained*

$$
  dS_A/dt = -S_A*(\beta_A*I_A + \beta_B*I_B + \beta_AB*I_{AB} + \beta_0*I + V_B) + S*V_A
$$ 

$-S_A*V_B$ -\> Remove individuals that get vaccine B (they move into the
S_AB population)\
$S*V_A$ -\> Add unvaccinated individuals that get vaccinated with
vaccine A\

$$
   dS_B/dt = -S_B*(\beta_A*I_A + \beta_B*I_B + \beta_AB*I_{AB} + \beta_0*I + V_A) + S*V_B
$$\
$-S_B*V_A$ -\> Remove individuals that get vaccine A (they move into the
S_AB population)\
$S*V_B$ -\> Add unvaccinated individuals that get vaccinated with
vaccine B\

$$
   dS_{AB}/dt = -S_{AB}*(\beta_A*I_A + \beta_B*I_B + \beta_{AB}*I_{AB} + \beta_0*I) 
   + S*V_{AB} + S_A*V_B + S_B*V_A
$$\
$S*V_{AB}$ -\> Add unvaccinated individuals that get both vaccines at
the same time\
$S_A*V_B$ -\> Add individuals already with vaccine A that get vaccine B\
$S_B*V_A$ -\> Add individuals already with vaccine B that get vaccine B\

### Infected

$$
   dI = S*(\beta_A*I_A + \beta_B*I_B + \beta_AB*I_{AB} + \beta_0*I) - 
   I*(c + \gamma_0 + V_A + V_B + V_{AB})
$$\
Breakdown\
$S*\beta_A*I_A$ - Add susceptible individuals that get infected by
individuals with vaccine A\
$S*\beta_B*I_B$ - Add susceptible individuals that get infected by
individuals with vaccine B\
$S*\beta_B*I_{AB}$ - Add susceptible individuals that get infected by
individuals with vaccine A&B\
$S*\beta_0*I$ - Add susceptible individuals that get infected by
unvaccinated individuals

$-I*c$ - remove individuals that are culled\
$-I*\gamma_0$ - remove individuals that recover\
$-I*V_A$ - remove unvax individuals that are infected and get vaccine A\
$-I*V_B$ - remove unvax individuals that are infected and get vaccine B\
$-I*V_{AB}$ - remove individuals that are infected and get vaccine A&B\

$$
   dI_A <- S_A*(\beta_A*I_A + \beta_B*I_B + \beta_AB*I_{AB} + \beta_0*I) - 
      I_A*(c + \gamma_A + V_B) + I*V_A
$$\
$I*V_A$ - Add uninfected, infected individuals that get vaccine A

$$
    dI_B <- S_B*(\beta_A*I_A + \beta_B*I_B + \beta_{AB}*I_{AB} + \beta_0*I) - 
      I_B*(\gamma_B + c + V_A) + I*V_B
$$\
$I*V_A$ - Add uninfected, infected individuals that get vaccine B

$$
    dI_{AB} <- S_{AB}*(\beta_A*I_A + \beta_B*I_B + \beta_{AB}*I_{AB} + \beta_0*I) - 
      I_{AB}*(\gamma_{AB} + c) + I*V_{AB} + I_A*V_B + I_B*V_A
$$\
$I*V_{AB}$ - Add uninfected, infected individuals that get vaccine A&B\
$I_A*V_B$ - Add vaccine A, infected individuals that get vaccine B\
$I_B*V_A$ - Add vaccine B, infected individuals that get vaccine A\

### Recovery

$$
\begin{aligned}
dR = I*\gamma_0 \\
dR_A = I_A*\gamma_A \\
dR_B = I_B*\gamma_B \\
dR_{AB} = I_{AB}*\gamma_{AB} \\
\end{aligned}
$$ $I*\gamma_0$ - unvaccinated individuals that recover\
$I_A*\gamma_A$ - vaccinated A individuals that recover\
$I_B*\gamma_B$ - vaccianted B individuals that recover\
$I_AB*\gamma_AB$ - vaccinated AB individuals that recover\


## CODE

```{r, echo = FALSE}
library(deSolve)
library(tidyverse)
```

```{r}
#### Set up ####

# define parameters
beta_0 <- 0  # Unvaccinated transmission rate
beta_A <- 0  # Vaccinated A transmission rate
beta_B <- 0.6  # Vaccinated B transmission rate
beta_AB <- 0 # Transmission rate of individual with AB vaccine

gamma_0 <- 0.15 # Unvaccinated recovery rate
gamma_A <- 0.15 # Recovery rate for vaccine A
gamma_B <- 0.15   # Recovery rate for vaccine B
gamma_AB <- 0.15  # Recovery rate for both vaccines

c <- 0  # Rate of culling

V_A <- 0 # vaccine rate of vaccine A
V_B <- 1 # vaccine rate of vaccine B
V_AB <- 0 # rate of getting vaccinated by both A and B


#### Model ####

# put parameter values into vector params
params <- c(beta_0 = beta_0, beta_A = beta_A, beta_B = beta_B, beta_AB = beta_AB,
            gamma_0 = gamma_0, gamma_A = gamma_A, gamma_B = gamma_B, gamma_AB = gamma_AB,
            c = c, V_A = V_A, V_B = V_B, V_AB = V_AB)

# define initial conditions
initial_S <- 0
initial_I <- 0
initial_R <- 0
initial_S_A <- 100
initial_S_B <- 0
initial_S_AB <- 0
initial_I_A <- 0
initial_I_B <- 1
initial_I_AB <- 0
initial_R_A <- 0
initial_R_B <- 0
initial_R_AB <-0

state <- c(S = initial_S, S_A = initial_S_A, S_B = initial_S_B, S_AB = initial_S_AB, 
           I = initial_I, I_A = initial_I_A, I_B = initial_I_B, I_AB = initial_I_AB,
           R = initial_R, R_A = initial_R_A, R_B = initial_R_B, R_AB = initial_R_AB) 

# define time
times <- seq(0, 100, 0.001)

# define the model!
model <- function(time, state, params){
  with(as.list(c(state,params)),{
    
    # Suseceptible 
    dS <- -S*(beta_A*I_A + beta_B*I_B + beta_AB*I_AB + beta_0*I + V_A + V_B + V_AB)
    dS_A <- -S_A*(beta_A*I_A + beta_B*I_B + beta_AB*I_AB + beta_0*I + V_B) + S*V_A
    dS_B <- -S_B*(beta_A*I_A + beta_B*I_B + beta_AB*I_AB + beta_0*I + V_A) + S*V_B
    dS_AB <- -S_AB*(beta_A*I_A + beta_B*I_B + beta_AB*I_AB + beta_0*I) + S*V_AB + S_A*V_B + S_B*V_A
    
    # Infected
    dI <- S*(beta_A*I_A + beta_B*I_B + beta_AB*I_AB + beta_0*I) - I*(c + gamma_0 + V_A + V_B + V_AB)
    dI_A <- S_A*(beta_A*I_A + beta_B*I_B + beta_AB*I_AB + beta_0*I) - I_A*(gamma_A + c + V_B) + I*V_A
    dI_B <- S_B*(beta_A*I_A + beta_B*I_B + beta_AB*I_AB + beta_0*I) - I_B*(gamma_B + c + V_A) + I*V_B
    dI_AB <- S_AB*(beta_A*I_A + beta_B*I_B + beta_AB*I_AB + beta_0*I) - I_AB*(gamma_AB + c) + I*V_AB + I_A*V_B + I_B*V_A
    
    # Recovery
    dR <- I*gamma_0 - (V_A+V_B+V_AB)*R
    dR_A <- I_A*gamma_A - (V_B+V_AB)*R_A + V_A*R
    dR_B <- I_B*gamma_B - V_A*R_B + V_B*R
    dR_AB <- I_AB*gamma_AB + V_AB*R + V_A*R_B + V_B*R_A
    
    return(list(c(dS, dS_A, dS_B, dS_AB, dI, dI_A, dI_B, dI_AB, dR, dR_A, dR_B, dR_AB)))
  })
}

# numerically integrate equations!
out <- as.data.frame(ode(state, times, model, params))

# plot

# plots everything
out |> select(time, I, I_A, I_B, I_AB) |> pivot_longer(! time) %>%
  ggplot(aes(x = time, y = value, color = name)) +
  geom_line()

# plots infection
matplot(out[, "time"], out[, c("I","I_A", "I_B", "I_AB")], type = "l", 
        col = c("purple","red", "blue", "green"), xlab = "Time", ylab = "Infected Individuals",
        main = "Disease Spread with Vaccination and Culling")
legend("topright", legend = c("Unvaccinated","Vaccine A", "Vaccine B", "Vaccine A and B"),
       col = c("purple","red", "blue", "green"), lty = c(1,1,1,1))


# plots susceptible
plot(x = out$time, y = out$S, col = "purple", type = "l")
# lines(x = out$time, y = out$S_A, col = "red")
lines(x = out$time, y = out$S_B, col = "blue")
lines(x = out$time, y = out$S_AB, col = "green")


# NEW
for (beta_AB in seq(0,1,0.1)){
  params$beta_AB <- beta_AB
  out <- as.data.frame(ode(state, times, model, params))
  out$I_AB[which.max(out$I_AB)]
}
#^ What this is:
# basically we are simulating different beta_AB values and finding the highest number of vaccinations and we can plot it
# Why? We do this to confirm expectations -> if beta_AB is increasing we expect I_AB to increase as well ig
# idk why are we doing this...
```

Next Steps:

-   verify the code is working

-   Verify the accuracy of the code

-   Get interpretable results
